---
title: "Predicting Stroke in Healthcare Patients"
author: '"Saba_Alemayehu"'
date: "6/20/2022"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(caret)
library(e1071)
library(Hmisc)
library(corrplot)
library(plyr)
library(pROC)
```

# Importing the data
```{r}
stroke_CSV<- read.csv("dataset-stroke-data.csv")
```

```{r}
table(stroke_CSV$stroke)
```

# encoding the variables into numeric and factors
```{r}
stroke_CSV$bmi <- as.numeric(stroke_CSV$bmi)
stroke_CSV$gender <- as.factor(stroke_CSV$gender)
stroke_CSV$work_type<-as.factor(stroke_CSV$work_type)
stroke_CSV$ever_married <- as.factor(stroke_CSV$ever_married)
stroke_CSV$Residence_type <- as.factor(stroke_CSV$Residence_type)
stroke_CSV$smoking_status <- as.factor(stroke_CSV$smoking_status)

stroke_CSV$stroke <- ifelse(stroke_CSV$stroke == 1, "yes", "no")

stroke_CSV$stroke<-as.factor(stroke_CSV$stroke)
table(stroke_CSV$stroke)
```

# Data Visualization
```{r}
# histogram of numeric variables
par(mfrow=c(1,3))
hist( stroke_CSV$age ) 
hist( stroke_CSV$bmi )
hist( stroke_CSV$avg_glucose_level )
```
# convert factors into dummy variables

```{r}
#create dummy variable 
dummy_varv02<-dummyVars(~gender+ever_married + work_type+ Residence_type+smoking_status, data=stroke_CSV,fullRank = T)
dummy_df1<-data.frame(predict(dummy_varv02,stroke_CSV))
head(dummy_df1)
# combine origional data and dummy data frame
combinedDf<-cbind(dummy_df1,stroke_CSV)
head(combinedDf)
# remove the redundunt column
stroke_dm <-dplyr::select(combinedDf, -c('gender', 
                                         'ever_married',
                                         'work_type',
                                         'Residence_type', 
                                         'smoking_status'))
                                         
                                            
                                           
                                           
                                           
#head(stroke_dm)
str(stroke_dm)
```

# remove unwanted columns
```{r}
stroke_dm = subset(stroke_dm, select = -c(id) )
#stroke_dm$stroke<-as.factor(stroke_dm$stroke)
str(stroke_dm)
```

```{r}
#head(stroke_dm)
table(stroke_dm$stroke)
```


# Data Spliting and data preprocessing
```{r}
pre_heni = subset(stroke_dm, select = -c(stroke) )
pre_heni
res_heni<-stroke_dm$stroke

set.seed(100)
training_rowSt1<-createDataPartition(res_heni, p=0.8, list=FALSE)

strok_prTrainX2<-pre_heni[training_rowSt1,]
strok_prTestX1<-pre_heni[-training_rowSt1,]

stroke_reTrY1<-res_heni[training_rowSt1]
stroke_reTesY1<-res_heni[-training_rowSt1]
str(stroke_reTrY1)
```

# Impute the missing value
```{r}
trainimpu1<-preProcess(strok_prTrainX2,"knnImpute")

strokeTrprX<-predict(trainimpu1,strok_prTrainX2)


strokeTeprX<-predict(trainimpu1,strok_prTestX1)
#str(strokeTeprX)
```


# Develop a model
```{r}
#library(caret)
#library(pROC)
#Logistic Regression
ctrl<-trainControl(method="LGOCV",
                   summaryFunction=twoClassSummary,
                   classProbs = TRUE,
                   savePredictions = TRUE,
                   sampling = "up")
                
set.seed(300)
lrSfit<-train(x=strokeTrprX, y=stroke_reTrY1,
             method = "glm",
             metric="ROC",
             preProcess = c("center","scale"),
             trControl=ctrl)
lrSfit
```

# Linear Discriminant Analysis:

```{r}
set.seed(300)
lda.classifier = train(x=strokeTrprX, 
                       y=stroke_reTrY1,
                       method = "lda",
                       preProc = c("center", "scale"),
                       metric = "ROC",
                       trControl = ctrl)
lda.classifier
```

# Penalized Methods
```{r}
glmnGrid = expand.grid(.alpha = c(0, 0.1, 0.2, 0.4, 0.6, 0.8, 1.0), .lambda = seq(0.01, 0.2, length=40))
set.seed(300)
glmnet.classifier = train(x=strokeTrprX,
                          y=stroke_reTrY1,
                          method = "glmnet",
                          tuneGrid = glmnGrid,
                          preProc = c("center", "scale"),
                          metric = "ROC",
                          trControl = ctrl)
glmnet.classifier
```

# Nearest shrunken Centroids:
```{r}
# Nearest shrunken Centroids:
nscGrid = expand.grid(.threshold = 0:25)
set.seed(300)
nsc.classifier = train(x=strokeTrprX,
                       y=stroke_reTrY1,
                       method = "pam",
                       tuneGrid = nscGrid,
                       preProc = c("center", "scale"),
                       metric = "ROC",
                       trControl = ctrl)
nsc.classifier

```

# k-nearest model
```{r}
set.seed(300)
knnModel = train(x=strokeTrprX,
                 y=stroke_reTrY1, 
                 method="knn", 
                 preProc = c("center", "scale"), 
                 tuneLength=10,
                 trControl = ctrl)
knnModel
```


# A Neural Network model

```{r}

nnGrid = expand.grid( .decay=c(0,0.01,0.1), .size=1:10, .bag=FALSE )
set.seed(300)
nnetModel = train(x=strokeTrprX,
                  y=stroke_reTrY1, 
                  method="nnet",
                  metric = "ROC",
                  preProc = c("center", "scale"),
                  linout=FALSE,
                  trace=FALSE,
                  MaxNWts=10 *(ncol(strokeTrprX)+1) + 10 + 1, maxit=500,trControl = ctrl)
nnetModel

```

# MARS model
```{r}
marsGrid = expand.grid(.degree=1, .nprune=2:10)
set.seed(300)
marsModel = train(x=strokeTrprX, 
                  y=stroke_reTrY1,
                  method="earth",
                  metric = "ROC",
                  preProc = c("center", "scale"),
                  tuneGrid=marsGrid,
                  trControl = ctrl)

marsModel  
str(stroke_reTrY1)

```

# A Support Vector Machine (SVM)

```{r}

set.seed(300)
svmModel = train(x=strokeTrprX,
                 y=stroke_reTrY1,
                 method="svmRadial",
                 ppreProc = c("center", "scale"),
                 tuneLength=10)
                 

svmModel

```

# Random Forest 
```{r}
library(randomForest)
rfModel = randomForest( stroke_reTrY1 ~ ., data=strokeTrprX,metric = "ROC", importance=TRUE, ntree=1000,trControl = ctrl)
rfModel
```



