---
title: "Predicting Stroke"
author: "Stephen Kuc"
date: '2022-06-14'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Importing dataset and necessary libraries
```{r}
stroke <- read.csv("c:/Users/steph/OneDrive/Documents/USD/ADS503/healthcare-dataset-stroke-data.csv")
```
```{r}
library(caret) # for training models
library(e1071) 
library(Hmisc)
library(corrplot)
library(plyr)
library(pROC)
```
```{r}
str(stroke)
```
Many of the categorical variables are characters -- we will need to change those to factors. 
```{r}
dim(stroke)
```

There are 5110 observations, with 12 features, including the target variable.
```{r}
# changing datatypes to what they should be
stroke$hypertension <- as.factor(stroke$hypertension)
stroke$heart_disease <- as.factor(stroke$heart_disease)
stroke$gender <- as.factor(stroke$gender)
stroke$ever_married <- as.factor(stroke$ever_married)
stroke$work_type <- as.factor(stroke$work_type)
stroke$Residence_type <- as.factor(stroke$Residence_type)
stroke$smoking_status <- as.factor(stroke$smoking_status)
stroke$bmi <- as.numeric(stroke$bmi)
stroke$stroke <- as.factor(stroke$stroke)
```

```{r}
# checking nulls
colSums(is.na(stroke))
```
There are 201 nulls in BMI. 
```{r}
summary(stroke)
```
1 "other" gender. 1544 "unknown" smoker status. 201 nulls in BMI. Work type "Private" means what? Any cutoff for minimum Age?
Target variable Stroke seems pretty imbalanced.
```{r}
# plots for all features
par(mar = c(2,2,2,2))
layout.matrix <- matrix(c(1,4,5,6,2,7,8,9,3,10,0,0),nrow = 4, ncol = 3)

layout(mat = layout.matrix,
       heights = c(4, 4, 4, 4), 
       widths = c(3, 3, 3))
# histogram for numerical features
hist(stroke$age, cex.main = .5, cex.axis = .5)
hist(stroke$avg_glucose_level, cex.main = .5)
hist(stroke$bmi, cex.main = .5)

# bar charts for categorical
countGen <- table(stroke$stroke,stroke$gender)
barplot(countGen, main = "Stroke distribution by Gender", legend = rownames(countGen), cex.lab = .5, cex.main = .5)

countHyp <- table(stroke$stroke, stroke$hypertension)
barplot(countHyp, main = "Stroke distribution by Hypertension", legend = rownames(countHyp), cex.lab= .5, cex.main = .5)

countHd <- table(stroke$stroke,stroke$heart_disease)
barplot(countHd, main = "Stroke distribution by Heart Disease", legend = rownames(countHd), cex.lab = .5, cex.main = .5)

countMarried <- table(stroke$stroke, stroke$ever_married)
barplot(countMarried, main = "Stroke distribution by if ever married", legend = rownames(countMarried), cex.lab = .5, cex.main = .5)

countWork <- table(stroke$stroke, stroke$work_type)
barplot(countWork, main = "Stroke distribution by Work Type", legend = rownames(countWork), cex.lab = .5, cex.main = .5, cex.axis = .5)

countRes <- table(stroke$stroke, stroke$Residence_type)
barplot(countRes, main = "Stroke distribution by Residence Type", legend = rownames(countRes), cex.lab = .5, cex.main = .5)

countSmoke <- table(stroke$stroke, stroke$smoking_status)
barplot(countSmoke, main = "Stroke distribution by Smoking Status", legend = rownames(countSmoke), cex.lab = .5, cex.main = .5, cex.axis = .5)
```
```{r}
# glucose levels look skewed slightly, as does bmi numbers
# let's check for skewness
skewness(stroke$avg_glucose_level)
skewness(stroke$bmi) ## need to get rid of nulls to see skewness metric
skewness(stroke$age)
```
```{r}
# let's investigate the numeric variables further
ggplot(stroke, aes(x=age,color=stroke)) + geom_histogram(bins = 20)

ggplot(stroke, aes(x=avg_glucose_level,color=stroke)) + geom_histogram(bins = 20)

ggplot(stroke, aes(x=bmi,color=stroke)) + geom_histogram(bins = 20)
```
